Child Poverty Interventions

### 2024-05-16 minos/outcomes/child_poverty_reduction_vis.Rmd
Tracking the ChildPovReduction plotting notebooks

require(tidyverse)
require(ggplot2)
require(ggridges) # for densities over time plots. 
require(knitr)
require(here)
require(gghighlight)

@@

```{r "imports", include=FALSE}
source(here::here('minos', 'utils_datain.R'))
#source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

```{r, include=FALSE}
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
int.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionRANDOM', drop.dead = TRUE)
int2.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionSUSTAIN', drop.dead = TRUE)
#base.dat <- read_agg_subset_batch_out(out.path, 'baseline', drop.dead = TRUE)
#int1.dat <- read_agg_subset_batch_out(out.path, 'ChildPovertyReduction', drop.dead = TRUE)
```

@@

We can calculate the relative poverty threshold for our starting year of 2020:

```{r}
base.2020 <- base.dat %>%
  filter(time == 2020)

mean.hh_income <- weighted.mean(base.2020$hh_income, w = base.2020$weight)
rel.pov.threshold <- mean.hh_income * 0.6

print(paste0("Relative poverty threshold in 2020: ", rel.pov.threshold))
rm(base.2020, mean.hh_income, rel.pov.threshold)
```

@@

## Random Uplift

@@

@@

#### Full Sample

@@

### True Treatment on Treated

Instead of looking at the difference in means

## Sustained Uplift

This intervention is similar to the random uplift intervention with one difference. In the first year that the intervention is applied, households are randomly chosen for uplift until the correct number of children for that year are uplifted. The difference comes in subsequent years. From year 2 onwards, households that have been previously intervened on will once again be uplifted to above the poverty line (if they are once again below). This is the meaning of the sustained intervention, where families are chosen at random and then maintained above the poverty line, instead of picking a random subset of households under the poverty line as in the random intervention.

@@

>>>>

library(tidyverse)
library(ggplot2)
library(ggridges) # for densities over time plots. 
library(knitr)
library(here)
library(gghighlight)

@@

```{r "imports", include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

```{r, include=FALSE}
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE,
                                    drop.zero.weight = TRUE)
int.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionRELATIVE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)
int2.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionABSOLUTE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)

#base.dat <- read_agg_subset_batch_out(out.path, 'baseline', drop.dead = TRUE)
#int1.dat <- read_agg_subset_batch_out(out.path, 'ChildPovertyReduction', drop.dead = TRUE)
```

@@

We can calculate the relative poverty threshold for our starting year of 2021:

```{r}
base.2021 <- base.dat %>%
  filter(time == 2021)

mean.hh_income <- weighted.mean(base.2021$hh_income, w = base.2021$weight)
rel.pov.threshold <- mean.hh_income * 0.6

print(paste0("Relative poverty threshold in 2020: ", rel.pov.threshold))
rm(base.2021, mean.hh_income, rel.pov.threshold)
```

@@

## Intervention

@@

hh_boosted <- int.dat %>%
    select(hidp, time, income_boosted, boost_amount, nkids) %>%
    filter(income_boosted == 'True') %>%
    group_by(time) %>%
    summarise(n_boosted = length(unique(hidp)),
              total_boost = sum(boost_amount),
              kids_boosted = sum(nkids))

@@

#### Full Population

@@

## Absolute Poverty

@@

# Diagnostics

## Spaghetti

```{r}
base.spag <- base.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'baseline')
int.spag <- int.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'relative')
int2.spag <- int2.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'absolute')

# spaghetti_plot(base.spag, 'hh_income',
#                save = FALSE,
#                save.path = '')
# 
# spaghetti_plot(int.spag, 'hh_income',
#                save = FALSE,
#                save.path = '')
# 
# spaghetti_plot(int2.spag, 'hh_income',
#                save = FALSE,
#                save.path = '')

density_ridges(base.spag, 'hh_income',
               save=FALSE, 
               save.path='')

density_ridges(int.spag, 'hh_income',
               save=FALSE, 
               save.path='')

density_ridges(int2.spag, 'hh_income',
               save=FALSE, 
               save.path='')
```

dens.1 <- rbind(base.spag, int.spag)
dens.2 <- rbind(base.spag, int2.spag)

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int.dat, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Relative Poverty Intervention')

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int2.dat, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Absolute Poverty Intervention')
